---
title: "lesson3_exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries and data

```{r}
library(ggplot2)
data(diamonds)
```

## Inspecting dataset

```{r}
?diamonds

str(diamonds)

levels(diamonds$color)

summary(diamonds)

```
```{r}
qplot(x = price, data = diamonds)
```

O histograma é distorcido à direita. Os valores estão concentrados na mediana (2401), os valores extremos à direita estão puxando a média para cima (3933)

## Diamond counts

```{r}
nrow(subset(diamonds,price < 500))
nrow(subset(diamonds,price < 250))
nrow(subset(diamonds,price >= 15000))
```

# Cheap diamonds

```{r}
qplot(x = price, data = diamonds, binwidth = 50) +
  scale_x_continuous(limits = c(250, 1500), breaks = seq(250, 1500, 50))

#ggsave('priceHistogram.png')
```

## Histogram diamond prices by cut

```{r}
qplot(x = price, data = diamonds) +
  facet_wrap(~cut, ncol = 3)
```

## Price summary by cut

```{r}
by(diamonds$price, diamonds$cut, summary)
```

## Changing histogram scales with facet_wrap

```{r}
qplot(x = price, data = diamonds) +
  facet_wrap(~cut, scales="free_y", ncol = 3)
```

## Histogram of price per carat

```{r}
qplot(x = price / carat, data = diamonds, binwidth = .05) +
  scale_x_log10() +
  facet_wrap(~cut, scales="free_y", ncol = 3)
```

## Box plot investigation

```{r}
qplot(x = clarity, y = price,
  data = diamonds, geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 7500))
```
```{r}
by(diamonds$price, diamonds$clarity, summary)
```

## Interquartile range

```{r}
by(diamonds$price, diamonds$color, summary)
```
```{r}
IQR(subset(diamonds, color == "D")$price)
IQR(subset(diamonds, color == "J")$price)
```

## Box plot price per carat

```{r}
qplot(x = color, y = price / carat,
  data = diamonds, geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 6000))
```

## Frequency polygon by carat

```{r}
qplot(x = carat, data = diamonds, geom = 'freqpoly', binwidth = .1) + 
  scale_x_continuous(breaks = seq(0, 6, .1))
```

```{r}
margin.table(table(diamonds$carat), 1)
```

