Lesson 4
========================================================

***

```{r Pseudo-Facebook User Data}
pf <- read.csv('https://s3.amazonaws.com/udacity-hosted-downloads/ud651/pseudo_facebook.tsv', sep = '\t')
```

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
qplot(x = age, y = friend_count, data = pf);
```

***

#### What are some things that you notice right away?
Response:

Younger people have a higher friend count, also "very old people", which are probably fake accounts based on previous analysis.

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
summary(pf$age)

ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point() + 
  xlim(13, 90)
```

***

### Overplotting
Notes:

To make points visibility better, we can add jitter (each point will be slighly offset) and transparancy to the points (so areas with higher concentration of data will appear darker)

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_jitter(alpha = 1/20) + #1/20 means that 20 points are necessary to make a "whole" point
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:

Spots where there's a higher level of data concentration are now more evident, outliers are not as visible.

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20) + 
  xlim(13, 90) + 
  coord_trans(y = "sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(height = 0)) + 
  xlim(13, 90) + 
  coord_trans(y = "sqrt")
```

#### What do you notice?

The lower layers distribution are better viewed.

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_point(alpha = 1/20) + 
  xlim(13, 90) + 
  coord_trans(y = "sqrt")
```

There's a good concentration of users that didn't initiated any friendships.

```{r}
ggplot(aes(x = age, y = friend_count), data = subset(pf, friendships_initiated == 0)) + 
  geom_point(alpha = 1/20) + 
  xlim(13, 90) + 
  coord_trans(y = "sqrt")
```
***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
#install.packages('dplyr')

library(dplyr)

#Without function chaining
age_groups <- group_by(pf, age)

pf.fc_by_age <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())

pf.fc_by_age <- arrange(pf.fc_by_age, age)

head(pf.fc_by_age)

#With function chaining
pf.fc_by_age <- pf %>% 
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)

head(pf.fc_by_age)

```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + 
  geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20, 
             position = position_jitter(height = 0), 
             color = 'orange') + 
  xlim(13, 90) +
  coord_trans(y = "sqrt") + 
  geom_line(stat = 'summary', fun.y = mean) + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .1), linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .9), linetype = 2, color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .5), color = 'blue')
  
```

#### What are some of your observations of the plot?
Response:

All the metrics rise when you look at the age 69 and the one at the right end of the chart. The number of friends seems to stabilize between age 30 and 70 (excluding 69). Having more than 1000 friends is very rare, as 90% of the users are bellow this threshold.
***

#### Zooming in

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20, 
             position = position_jitter(height = 0), 
             color = 'orange') + 
  coord_cartesian(xlim = c(13, 70), ylim = c(0, 1000)) + 
  geom_line(stat = 'summary', fun.y = mean) + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .1), linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .9), linetype = 2, color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .5), color = 'blue')
```


### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
?cor.test
cor.test(pf$age, pf$friend_count, method = 'pearson')

#Using with
with(pf, cor.test(age, friend_count, method = 'pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

-0.02740737 

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(height = 0)) +
  xlim(0, 2500) + 
  ylim(0, 2500)
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(height = 0)) +
  xlim(0, quantile(pf$www_likes_received, .95)) + 
  ylim(0, quantile(pf$likes_received, .95)) + 
  geom_smooth(method = 'lm', color = 'red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf, cor.test(www_likes_received, likes_received))
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)

data(Mitchell)

```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
  geom_point() +
  geom_smooth(method = 'lm', color = 'red')
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

It's about zero

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
range(Mitchell$Month)

ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
  geom_point() +
  scale_x_continuous(breaks = seq(0, 203, 12))
```

```{r}
ggplot(aes(x=(Month%%12),y=Temp),data=Mitchell)+ 
  geom_point()
```


***

### A New Perspective

What do you notice?
Response:

A sine wave starts to appear

Watch the solution video and check out the Instructor Notes!
Notes:

***

```{r}
#install.packages('energy')
library(energy)

x <- seq(0, 4*pi, pi/20)
y <- cos(x)
qplot(x = x, y = y)
dcor.ttest(x, y)
```


### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- pf$age + ((12 - pf$dob_month) / 12)
```

***

### Age with Months Means

```{r Age with Months Means}
pf.fc_by_age_months <- pf %>% 
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months)
```

Programming Assignment
```{r Programming Assignment}
ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) + 
  geom_line()
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
#install.packages('gridExtra')
library(gridExtra)

p1 <- ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) + 
  geom_line() + 
  geom_smooth()

p2 <- ggplot(aes(x = age, y = friend_count_mean), data = subset(pf.fc_by_age, age < 71)) + 
  geom_line() + 
  geom_smooth()

p3 <- ggplot(aes(x = round(age / 5) * 5, y = friend_count), data = subset(pf, age < 71)) + 
  geom_line(stat = 'summary', fun.y = mean)

grid.arrange(p1, p2, p3, ncol = 1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

