Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
#install.packages('ggplot2')
library(ggplot2)

pf <- read.csv('https://s3.amazonaws.com/udacity-hosted-downloads/ud651/pseudo_facebook.tsv', sep = '\t')

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() + 
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```
```{r}
library(dplyr)

pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  arrange(age, gender) %>%
  ungroup()

head(pf.fc_by_age_gender)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count, color = gender),
       data = pf.fc_by_age_gender) + 
  geom_line()
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, 
                                  age ~ gender, 
                                  value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)
```

### Same solution using "tidyr" instear
```{r}
#install.packages("tidyr")
library(tidyr) 
spread(subset(pf.fc_by_age_gender, select = c('gender', 'age', 'median_friend_count')), gender, median_friend_count)
```

***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = female / male),
       data = pf.fc_by_age_gender.wide) + 
  geom_line() + 
  geom_hline(yintercept = 1, alpha = .3, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- 2014 - ceiling(pf$tenure / 365)

summary(pf$year_joined)

table(pf$year_joined)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket, useNA = 'ifany')

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) + 
  geom_line(color = 'black', linetype = 2, stat = 'summary', fun.y = mean)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure > 0), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?
0.2205

What is the maximum friend rate?
417.0000

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, !is.na(year_joined.bucket) & tenure > 0)) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) 
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}
#install.packages('gridExtra')
library(gridExtra)

p1 <- ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

p2 <- ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p3 <- ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p4 <- ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

grid.arrange(p1, p2, p3, p4, ncol = 1)
```

### Using geom_smooth

```{r}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))
```


***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')

str(yo)

yo$id <- factor(yo$id)

str(yo)

ggplot(aes(x = price), data = yo) + 
  geom_histogram(binwidth = 10)
```

There are several empty spots in the histogram, prices are not continous. After adjusting the bin width, the histogram became distorted to the left but it is very biased.

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo$price)
length(unique(yo$price))
table(yo$price)

#?transform
names(yo)
yo <- transform(yo, all.purchases = (strawberry + blueberry + pina.colada + plain + mixed.berry))

ggplot(aes(x = all.purchases), data = yo) + 
  geom_histogram(binwidth = 1)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x = time, y = price), data = yo) + 
  geom_jitter(alpha = 0.25, shape = 21, fill = I('#F79420'))
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)

sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price), data = subset(yo, id %in% sample.ids)) + 
  facet_wrap(~ id) +
  geom_line() + 
  geom_point(aes(size = all.purchases), pch = 1)
  
```
```{r}
set.seed(9137)

sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price), data = subset(yo, id %in% sample.ids)) + 
  facet_wrap(~ id) +
  geom_line() + 
  geom_point(aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
#install.packages('GGally')

library(GGally)

set.seed(1836)
pf_subset <- pf[, c(2:15)]

names(pf_subset)

#takes a long time to run
#ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])
```


***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("https://s3.amazonaws.com/udacity-hosted-downloads/ud651/nci.tsv")
colnames(nci) <- c(1:64)
```

Note on melt;

melt will convert columns and collapse them as rows

e.g. from:

|id|x|y|
|-|-|-|
|1|1|0|
|2|1|1|
|3|2|3|

to:

|id|variable|value|
|-|-|-|
|1|x|1|
|1|y|0|
|2|x|1|
|2|y|1|
|3|x|2|
|3|y|3|

```{r}
library(reshape2)

nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:
We learnt how to compare multiple variables incorporating shapes and colors to condify information along with techniques to reduce noise in charts. Also how to use scatter matrixes and heatmaps to view a large amount of data in a single visualization.

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

